Fast, Scalable Phrase-Based SMT Decoding

### 2016-02-09 acl.2016/acl2016.tex
acl paper

We will focus on two main optimizations that differs from the current Probing phrase-table implementation and that have application to other implementations. Firstly, we will look at the \textbf{caching} of often used translation rules. The default caching framework save the most recently looked up rules. However, this caching mechanism is actually slower than re-querying the phrase-table, Table~\ref{tab:phrase-table-cache}. 
\begin{table}[h]
\small
\begin{center}
\begin{tabular}{|l|c|c|} \hline
		& No cache	& Caching \\ \hline
Decoding time  	& 2032 		& 2302 \\ \hline
\end{tabular}
\end{center}
\caption{Decoding time (in sec with 32 threads) when using phrase-table cache}
\label{tab:phrase-table-cache}
\end{table}
We shall explore a simpler caching mechanism that creates the cache at the start of decoding and remain static thereafter.

Secondly, the Probing phrase-table use a simple \textbf{compression} algorithm to compress the target side of the translation rule. Compression was championed being the main factor for the speed of the Compact phrase-table but as we saw in Figure~\ref{fig:moses-phrase-tables-time}, it may come at a cost when used with higher number of cores. We shall therefore take the opposite approach to~\newcite{junczys_tsd_2012b} and explore the decoding speed gain by disabling compression.

@@

Overall though, the scalability is remarkably good, the decoder is able to make full use of all real and virtual cores. When all cores are saturated, the decoder is 16 times faster than single-threaded decoding, on this 16 real core server. It is also 4.5 times faster than Moses with a single-thread and 9.6 faster when all cores are used.

@@

We verify the results on a smaller French-English phrase-based model and test set container longer sentences. The speed gain is even greater than our main testing scenario, Figure~\ref{fig:fr-en-speed}. Our decoder is 5.4 times faster than Moses with a single-thread and 14.5 faster when all cores are saturated.
\begin{figure}[h]
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.4]{fr-en-speed.png}} 
\end{tabular}
\caption{Decoding speed for fr-en model}
\label{fig:fr-en-speed}
\end{figure}

>>>>

We will focus on two main optimizations that differs from the current Probing phrase-table implementation and that have application to other implementations. Firstly, we will look at the \textbf{caching} of often used translation rules. The default caching framework save the most recently looked up rules. However, this caching mechanism is actually slower than re-querying the phrase-table, Table~\ref{tab:phrase-table-cache}. 
\begin{table}[h]
\small
\begin{center}
\begin{tabular}{|l|c|c|} \hline
		& No cache	& Caching \\ \hline
Decoding time  	& 2032 		& 2302 (+13\%) \\ \hline
\end{tabular}
\end{center}
\caption{Decoding time (in sec with 32 threads) when using phrase-table cache}
\label{tab:phrase-table-cache}
\end{table}
We shall explore a simpler caching mechanism that creates the cache at the start of decoding and remain static thereafter.

Secondly, the Probing phrase-table use a simple \textbf{compression} algorithm to compress the target side of the translation rule. Compression was championed as the main reason behind the speed of the Compact phrase-table but as we saw in Figure~\ref{fig:moses-phrase-tables-time}, this comes at the cost of scalability to large number of threads. We shall therefore take the opposite approach to~\newcite{junczys_tsd_2012b} and explore optimizing decoding speed by disabling compression.

@@

Overall though, the scalability is remarkably good, the decoder is able to make full use of all real and virtual cores. When all 16 cores are saturated with 2 hyper-thread each, the decoder is 16 times faster than single-threaded decoding. It is also 4.5 times faster than Moses with a single-thread and 9.6 faster when all cores are used.

@@

We verify the results on a smaller French-English phrase-based model and test set containng longer sentences. The speed gain is even greater than our main testing scenario, Figure~\ref{fig:fr-en-speed}. Our decoder is 5.4 times faster than Moses with a single-thread and 14.5 faster when all cores are saturated.
\begin{figure}[h]
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.4]{fr-en-speed.png}} 
\end{tabular}
\caption{Decoding speed for fr-en model}
\label{fig:fr-en-speed}
\end{figure}
